---
title: "COVID-19 Research"
author: "Benjamin Buss"
date: "6/13/2020"
output: pdf_document
---


```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
```


### Introduction

*Disclaimer: I'm not an Epidemiologist in any way shape or form.*

Analysis of COVID-19 data as a project for an undergrad statistics class at Utah Valley University.


### Data and Methods

#### Data


```{r, include = F, warning = F}
corona_counties_url <- "https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv"
corona_counties <- read_csv(corona_counties_url)

census_data_url <- "https://raw.githubusercontent.com/BenjaminBuss/COVID19_Research/master/processed_data/census_data.csv"
census_data     <- read_csv(census_data_url)

historical_weather_data_url <- "https://raw.githubusercontent.com/BenjaminBuss/COVID19_Research/master/processed_data/historical_weather.csv"
historical_weather <- read_csv(historical_weather_data_url)

current_weather_data_url <- "https://raw.githubusercontent.com/BenjaminBuss/COVID19_Research/master/processed_data/current_weather.csv"
current_weather <- read_csv(current_weather_data_url)

geographic_area_url <- "https://raw.githubusercontent.com/BenjaminBuss/COVID19_Research/master/processed_data/geographic_area.csv"
geographic_area     <- read_csv(geographic_area_url)

stations_url <- "https://raw.githubusercontent.com/BenjaminBuss/COVID19_Research/master/processed_data/zipcode_station.csv"
stations <- read_csv(stations_url)

rm(corona_counties_url, census_data_url, historical_weather_data_url, current_weather_data_url, geographic_area_url, stations_url)
```

```{r}

current_weather <- current_weather %>% mutate(temp = ( value / 10 ) * ( 9 / 5 ) + 32 )

```

Using the census data we selected the 60 most populous counties in the United States. All other data was selected to match these counties. 


```{r, include = F, warning = F} 

census_data <- census_data %>% top_n(60, population)

corona_counties_pop <- inner_join(corona_counties, census_data, by = "fips")

population_density <- inner_join(census_data, geographic_area, by = "fips") %>%
  transmute(fips, density = population / land_size)

```

#### Methods

All analysis was done using R and was designed to be fully reproducible solely based on this notebook.

For each of the 60 most populous counties we started by selecting a 29 period starting on the first day they observed at least one hundred cases of COVID-19. A growth factor was computed by taking that days cases divided by the previous days cases. Then a geometric mean was created from the 28 growth factors.

$$Growth Factor = ( \prod_{n = 2}^{28} \frac{Cases_{n day}}{Cases_{n-1 day}} )^{\frac{1}{n-1}} $$

```{r}

corona_counties_cut <- corona_counties_pop %>% filter(cases >= 100) %>% group_by(fips) %>% 
    arrange(date) %>% slice(1:29) 

growth_factors <- corona_counties_cut %>% mutate(case_growth = (cases / lag(cases))) %>% 
    filter(!is.na(case_growth))

geom_growth <- growth_factors %>% group_by(fips) %>% summarise(n = n(), geom_factor = prod(case_growth)^(1/n)     ) 

y <- inner_join(stations %>% select(station, fips), current_weather, by = "station") %>% 
    group_by(fips, date) %>% summarize(cur_temp = mean(temp))

historical_weather <- historical_weather %>% mutate(year = 2020) %>% unite(year, month, day, col = "date", sep = "-") %>% 
    mutate(date = as.Date(date))

yy <- inner_join(stations %>% select(station, fips), historical_weather, by = "station") %>% 
    group_by(fips, date) %>% summarize(his_temp = mean(temp))

zzz <- inner_join(corona_counties_cut, y, by = c("fips", "date")) %>%
    inner_join(., yy, by = c("fips", "date"))

counts <- count(zzz, fips)

```

For the historical and current weather data, we created a proportion of the probability that the temperature on a day in the time frame was between 40 and 50 degrees Fahrenheit.

```{r}

weather_proportions <- zzz %>%
  mutate(within = case_when(
    cur_temp >= 40 & cur_temp <= 50 ~ 1, TRUE ~ 0),
    his_within = case_when(
    his_temp >= 40 & his_temp <= 50 ~ 1, TRUE ~ 0)) %>% # Code column as 1 if within 40 & 50 degrees or 0 otherise
  group_by(fips) %>% summarise(n = n(), with = sum(within), his_with = sum(his_within)) %>%
  transmute(fips, current_proportion = with / n, historical_proportion = his_with / n)

```

```{r, include = F }

growth_data <- inner_join(geom_growth, weather_proportions, by = "fips") %>%
  inner_join(., population_density, by = "fips")

all_data <- growth_data %>% select(fips, geom_factor, density, current_proportion, historical_proportion)

```


Linear regressions


```{r}

simple_model <- lm(geom_factor ~ current_proportion, data = all_data)
added_model  <- lm(geom_factor ~ current_proportion + historical_proportion, data = all_data)
# https://rpubs.com/josevilardy/confounding ??
percentage_change <- (simple_model$coefficients[2] - added_model$coefficients[2])/simple_model$coefficients[2]*100; percentage_change
model <- lm(geom_factor ~ density + current_proportion + historical_proportion, data = all_data)
summary(model)
qqnorm(model$residuals)
cor(all_data$geom_factor, all_data$density)

summary(simple_model)
plot(simple_model)
```



### Results



### Discussion



### Limitations



### Acknowledgments



### References


#### Data Links:  


Case Counts: <https://www.nytimes.com/interactive/2020/us/coronavirus-us-cases.html>

  - <https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv>
  
  
Census Data: <https://www.census.gov/data/tables/time-series/demo/popest/2010s-counties-detail.html>

  - <https://raw.githubusercontent.com/BenjaminBuss/COVID_Project/master/census_data.csv>


Current Weather: <https://console.cloud.google.com/bigquery?p=bigquery-public-data&d=ghcn_d&page=dataset>

  - <https://raw.githubusercontent.com/BenjaminBuss/COVID_Project/master/current_weather_data.csv>


Historical Climate: <ftp://ftp.ncdc.noaa.gov/pub/data/normals/1981-2010/documentation/>

  - <https://raw.githubusercontent.com/BenjaminBuss/COVID_Project/master/historical_weather_data.csv>
  

Land Area: <https://www.census.gov/library/publications/2011/compendia/usa-counties-2011.html#LND>

  - <https://raw.githubusercontent.com/BenjaminBuss/COVID_Project/master/geographic_area.csv>
  

Stations and Zip codes <ftp://ftp.ncdc.noaa.gov/pub/data/normals/1981-2010/station-inventories/>

                       <https://www.kaggle.com/danofer/zipcodes-county-fips-crosswalk/version/1#>
                       
  - <https://raw.githubusercontent.com/BenjaminBuss/COVID_Project/master/stations.csv>




### Tables and Figures

```{r, warning=F}
us_map <- blscrapeR::county_map_data
us_growth <- corona_counties %>% filter(cases > 50) %>% group_by(fips) %>% arrange(date) %>%
  slice(1:22) %>%
  mutate(case_growth = (cases / lag(cases))) %>% filter(!is.na(case_growth))
us_geom <- us_growth %>% group_by(fips) %>% summarise(n = n(), geom_factor = prod(case_growth)^(1/n)) 
ggplot() +
  geom_map(data = us_map, map = us_map,
           aes(x = long, y = lat, map_id = id, group = group),
           fill = "#ffffff", color = "#0e0e0e", size = 0.15) +
  geom_map(data = us_geom, map = us_map, aes_string(map_id = "fips", fill = us_geom$geom_factor),
           color = "#0e0e0e", size = 0.15)
# https://cran.r-project.org/web/packages/blscrapeR/vignettes/Mapping_BLS_Data.html
```








